Quantifying Infants' Everyday Restrained Experiences in the Home Using Wearable Inertial Sensors

### 2025-10-02 manuscript_restrained_classification validation.qmd
updated figures and tables

```{r}
#| include: false
library(papaja)
library(knitr)
library(patchwork)
library(scales)
library(ggforce)
library(hms)
library(tidyverse)
library(lubridate)
library(rstatix)
library(janitor)
library(flextable)
library(effectsize)
```

@@

model_labels <- c("4_no_pos", "16_no_pos", "30_no_pos")
model_shapes <- c(5,1,2) %>% set_names(model_labels)
class_labels <- c("r", "u")
class_shapes <- c(5,1) %>% set_names(class_labels)
class_colors <- c("#0072B2","#E69F00") %>% set_names(class_labels)

indx <- c("balanced_accuracy","kappa", "sensitivity","pos_pred_value","f1")
indices <- c("Sensitivity", "Positive Predictive Value", "F1", "Prevalence Difference")

@@

```{r tbl-indexbetweenctgr}
#| apa-note: Age in months served as a continuous predictor and was centered within each age group; skill was a categorical predictor with non-sitters/non-walkers serving as the reference categories
#| tbl-cap: POSTPONE-Models' prediction performance for unrestrained periods compared to restrained periods
#| ft.align: left

# Compare between categories while controlling for models
lm_ctgr <- function(model_select){
  temp_ds <- matrix_comb %>% filter(model==model_select)
  list(
    sen = summary(lm(sensitivity ~ positive_class, data=temp_ds))$coefficients,
    ppv = summary(lm(pos_pred_value ~ positive_class, data=temp_ds))$coefficients,
    f1  = summary(lm(f1 ~ positive_class, data=temp_ds))$coefficients,
    prev_df = summary(lm(prevalence_df ~ positive_class, data=temp_ds))$coefficients
  )
}

#
t <- matrix_comb %>% 
  filter (model=="30_no_pos") %>% 
  select(id, session, positive_class, sensitivity) %>% 
  pivot_wider(names_from = positive_class, values_from = sensitivity) %>% 
  mutate(df = u-r)
summary(lm(df ~ 1, data = t))

#
tbl_ctgr_30 <- lm_ctgr("30_no_pos")%>% 
  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
  relocate("Indices", .before = "Estimate")%>% 
  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
         p = str_remove(p, "^0")) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  add_column(Term = rep(c("Intercept","Prediction of Unrestrained"), 4), .before = "Estimate") %>% 
  select(-SE)

tbl_ctgr_16 <- lm_ctgr("16_no_pos") %>% 
  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
  relocate("Indices", .before = "Estimate")%>% 
  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
         p = str_remove(p, "^0")) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  add_column(Term = rep(c("Int.","Unrest. vs Rest."), 4), .before = "Estimate") %>% 
  select(-SE)

tbl_ctgr_4 <- lm_ctgr("4_no_pos") %>% 
  map2_dfr(indices, ~ mutate(as.data.frame(.x), Indices = .y)) %>% 
  relocate("Indices", .before = "Estimate")%>% 
  rename(p = `Pr(>|t|)`, SE = `Std. Error`, t = `t value`) %>% 
  mutate(p = ifelse(p < .001,"<.001",round(p,3)),
         p = str_remove(p, "^0")) %>% 
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>% 
  add_column(Term = rep(c("Int.","Unrest. vs Rest."), 4), .before = "Estimate") %>% 
  select(-SE)

tbl_ctgr <- cbind(tbl_ctgr_4, tbl_ctgr_16 %>% select(-Indices,-Term) %>% rename(Estimate_16=Estimate,t_16=t,p_16=p)) %>% 
  cbind(tbl_ctgr_30 %>% select(-Indices, -Term) %>% rename(Estimate_30=Estimate,t_30=t,p_30=p))

tbl_flex_ctgr <- tbl_ctgr %>% 
  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>%
  set_header_labels(Estimate_16 = "Estimate", t_16 = "t", p_16 = "p",
                    Estimate_30 = "Estimate", t_30 = "t", p_30 = "p") %>%
  merge_v(j = "Indices") %>% fix_border_issues() %>% 
  hline(i = c(2,4,6,8)) %>% 
  padding(padding.top = 2, padding.bottom = 2) %>% 
  set_table_properties(layout = "autofit", width = .99) %>% 
  align(align = "center", part = "all") %>% 
  add_header_row(values = c("", "4s Model", "16s Model", "30s Model"), colwidths = c(2,3,3,3))
tbl_flex_ctgr
```

>>>>

```{r}
#| include: false
library(papaja)
library(knitr)
library(patchwork)
library(scales)
library(ggforce)
library(hms)
library(tidyverse)
library(lubridate)
library(rstatix)
library(janitor)
library(flextable)
library(effectsize)
library(ggbeeswarm)
```

@@

model_labels <- c("4_no_pos", "16_no_pos", "30_no_pos")
model_shapes <- c(5,1,2) %>% set_names(model_labels)
model_colors <- c("#0072B2","#E69F00", "#009E73") %>% set_names(c("4 s", "16 s", "30 s"))
class_labels <- c("r", "u")
class_shapes <- c(5,1) %>% set_names(class_labels)
# class_colors <- c("#0072B2","#E69F00") %>% set_names(class_labels)

indx <- c("balanced_accuracy","kappa", "sensitivity","pos_pred_value","f1")
per_class_indx <- c("sensitivity","pos_pred_value","f1")

@@

```{r tbl-cibtwctg}
#| apa-note: Compare Between Categories
#| tbl-cap: XX
#| ft.align: left
# For per-class metrics (i.e., sens, ppv, f1), calculate the CI of the metric difference btw classes (i.e., r and u)

# Function to bootstrap CI of gap for each model and each metric
bootstrap_gap <- function(df, metric, B = 2000) {
  df_wide <- df %>%
    select(session, id, positive_class, !!sym(metric)) %>%
    pivot_wider(names_from = positive_class, values_from = !!sym(metric)) %>%
    mutate(gap = r - u,
           id_session = paste(id, session, sep = "_"))

  sessions <- unique(df_wide$id_session)
  n <- length(sessions)
  res <- numeric(B)

  for (b in seq_len(B)) {
    sampled <- sample(sessions, size = n, replace = TRUE)
    df_b <- df_wide %>% filter(id_session %in% sampled)
    res[b] <- mean(df_b$gap, na.rm = TRUE)
  }

  ci <- quantile(res, c(0.025, 0.975), na.rm = TRUE)

  list(
    metric = metric,
    mean_gap = mean(df_wide$gap, na.rm = TRUE),
    ci = ci
  )
}

table_btw_ctg <- crossing(model = model_labels, metric = per_class_indx) %>%   # all combinations
  mutate(result = map2(model, metric, 
                       ~ bootstrap_gap(df = matrix_comb %>% filter(model == .x), 
                                       metric = .y))) %>%
  mutate(result = map(result, ~ tibble(
    mean_gap = .x$mean_gap,
    ci_low   = .x$ci[1],
    ci_high  = .x$ci[2]
  ))) %>% 
  unnest(result) %>% 
  mutate(model = factor(model, levels = c("4_no_pos","16_no_pos","30_no_pos")),
         metric = factor(metric, levels = c("sensitivity","pos_pred_value","f1"))) %>% 
  mutate(model = recode(model,
                        "4_no_pos" = "4 s",
                        "16_no_pos" = "16 s",
                        "30_no_pos" = "30 s"),
         metric = recode(metric,
                         "sensitivity" = "Sensitivity",
                         "pos_pred_value" = "Positive Predictive Value",
                         "f1" = "F1")) %>% 
  arrange(model, metric) %>% 
  rename("Window Length" = "model",
         "Metric" = "metric",
         "M (Rest.- Unrest.)"="mean_gap",
         "2.5% CI (Rest.- Unrest.)"="ci_low",
         "97.5% CI (Rest.- Unrest.)"="ci_high")

table_flex_btw_ctg <- table_btw_ctg %>% 
  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
  merge_v(j = "Window Length") %>% fix_border_issues() %>% 
  hline(i = c(3,6,9)) %>% 
  padding(padding.top = 2, padding.bottom = 2) %>% 
  set_table_properties(layout = "autofit", width = .99) %>% 
  align(align = "center", part = "all")
table_flex_btw_ctg
```

```{r fig-cibtwctg}
#| fig-cap: Compare between categories
#| fig-width: 7
#| fig-height: 5
#| apa-note: XX
p_ci <- ggplot(table_btw_ctg, 
       aes(x = Metric, 
           y = `M (Rest.- Unrest.)`, 
           ymin = `2.5% CI (Rest.- Unrest.)`, 
           ymax = `97.5% CI (Rest.- Unrest.)`, 
           color = `Window Length`)) +
  geom_pointrange(position = position_dodge(width = 0.5), size = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_color_manual(values=model_colors,labels=c("4 s","16 s", "30 s")) +
  scale_y_continuous(
    limits = c(-0.15, 0.1),  
    breaks = c(-0.15,-0.1, -0.05, 0, 0.05, 0.1) 
  )+
  theme_update() +
  labs(x = NULL,y = "CI of metric difference between restrained and unrestrained",
    color = "Window Length")

p_ci
# ggsave("figure/CI.png",width=7, height=5)
# ggsave("figure/CI.eps",width=7, height=5)

```{r tbl-prevalencegap}
#| apa-note: Diiference of Restraint Prevalence Between Model and Human
#| tbl-cap: XX
#| ft.align: left

df_prev <- matrix_comb %>% 
  filter(positive_class=="r") %>% 
  select(model, id, session, detection_prevalence, prevalence)

paired_t <- function(df){
  tt<-t.test(df$detection_prevalence, df$prevalence, paired=T)
  list(
    t = tt$statistic,
    p = tt$p.value
  )
}

table_prev <- df_prev %>% 
  group_by(model) %>% 
  summarise(actual_prev = mean(detection_prevalence),
            prev = mean(prevalence),
            .groups = "drop") %>% 
  mutate(result=map(model_labels, ~paired_t(df_prev %>% filter(model==.x))))%>% 
  mutate(result = map(result, ~tibble(
    t = .x$t,
    p = .x$p
  ))) %>% 
  unnest(result) %>% 
  mutate(model=recode(model,
                      "4_no_pos" = "4 s",
                       "16_no_pos" = "16 s",
                       "30_no_pos" = "30 s")) %>% 
  rename("Window Length" = "model",
         "Human Annotation Prevalence" = "actual_prev",
         "Model Prediction Prevalence" = "prev")

table_flex_prev <- table_prev %>% 
  flextable() %>% flextable::theme_apa() %>% line_spacing(part = "all") %>% 
  fix_border_issues() %>% 
  hline(i = c(1,2,3)) %>% 
  padding(padding.top = 2, padding.bottom = 2) %>% 
  set_table_properties(layout = "autofit", width = .99) %>% 
  align(align = "center", part = "all")
table_flex_prev

```{r fig-prevalence}
#| fig-cap: Prevalence Correlation
#| fig-width: 7
#| fig-height: 5
#| apa-note: XX
p_prev <- ggplot(data=matrix_comb %>% filter(positive_class=="r") %>%
                   mutate(model=recode(model,
                                       "4_no_pos" = "4 s",
                                       "16_no_pos" = "16 s",
                                       "30_no_pos" = "30 s")),
                 aes(x=prevalence, y=detection_prevalence, color=model))+
  geom_point(shape = 16,size = 5,alpha=.3) + 
  geom_abline(intercept = 0, slope = 1, linetype="dashed", color = "black") +
  geom_smooth(method = "lm", se = FALSE, size = 1) +  
  scale_color_manual(values=model_colors, labels=c("4 s", "16 s","30 s")) +
  scale_x_continuous(limits = c(0,1), breaks=c(0,.5,1)) +
  scale_y_continuous(limits = c(0,1), breaks=c(0,.5,1))+
  labs(x="Prevalence of restraint by human annotation", y="Prevalence of restraint by model prediction", color="Window Length")+
  theme_update()
p_prev

```{r fig-prevalencegap}
#| fig-cap: Prevalence Gap 
#| fig-width: 7
#| fig-height: 5
#| apa-note: XX
p_prev_gap <- ggplot(data=matrix_comb %>% filter(positive_class=="r") %>%
                   mutate(model=recode(model,
                                       "4_no_pos" = "4 s",
                                       "16_no_pos" = "16 s",
                                       "30_no_pos" = "30 s")),
                 aes(x=model, y=detection_prevalence-prevalence, color=model))+
  geom_violin(trim = FALSE, alpha = 0.4) +
  geom_quasirandom(aes(color = model),
                   dodge.width = 0.7, varwidth = TRUE, alpha = 0.6, size = 2) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white") + 
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  scale_color_manual(values=model_colors, labels=c("4 s", "16 s","30 s")) +
  labs(x="", y="Detection prevalence - true prevalence of restraint")+
  guides(color = "none")+
  theme_update()
p_prev_gap
```
